---
title: "Getting Started with Ryacas"
author: "Peter von Rohr"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Disclaimer
Test and experiments with using yacas through the package Ryacas.


# Resources
* Ryacas: https://github.com/r-cas/ryacas


# Installation
According to the Ryacas github page, the installation of the package is done via

```{r, eval=FALSE}
install.packages("Ryacas", dependencies = TRUE)
```


# Getting Started
According to http://r-cas.github.io/ryacas/articles/getting-started.html, we try to replicate the examples. 

```{r}
library(Ryacas)
```


# Low Level Interface

## Examples

```{r}
eq <- "x^2 + 4 + 2*x + 2*x"
yac_str(eq) # No task was given to yacas, so we simply get the same returned
```

Simplify the equation

```{r}
yac_str(paste0("Simplify(", eq, ")"))
```

Factor the equation

```{r}
yac_str(paste0("Factor(", eq, ")"))
```

The following command can be used to obtain an expression as a result

```{r}
(result_e <- yac_expr(paste0("Factor(", eq, ")")))
```

This can then be evaluated as follows

```{r}
eval(result_e, list(x = 2))
```

The TeX format is obtained via

```{r}
yac_str(paste0("TeXForm(Factor(", eq, "))"))
```


The helper function `y_fn()` can be used for convenience.

```{r}
y_fn(eq, "Factor")
```

Together with `yac_str` we get

```{r}
yac_str(y_fn(eq, "Factor"))
```

The magrittr pipe can also be used

```{r}
library(magrittr)
eq %>% y_fn("Factor") %>% yac_str()
```


# High Level Interface

## Examples

The high-level interface uses the function `ysym()`

```{r}
(eqy <- ysym(eq))
```

This can be converted into a R expression with

```{r}
as_r(eqy)
```

Factorisation can be done with 

```{r}
eqy %>% y_fn("Factor")
```


# Solving Equations
The Rosenbrock function is considered

```{r}
x <- ysym("x")
y <- ysym("y")
f <- (1 - x)^2 + 100*(y - x^2)^2
f
```

Showing it in TeX format

```{r, results='asis'}
cat("$$", tex(f), "$$\n")
```

The function can be visualised by 

```{r}
N <- 30
x <- seq(-1, 2, length=N)
y <- seq(-1, 2, length=N)
f_r <- as_r(f)
f_r
```

Creating a contour plot

```{r}
z <- outer(x, y, function(x, y) eval(f_r, list(x = x, y = y)))
levels <- c(0.001, .1, .3, 1:5, 10, 20, 30, 40, 50, 60, 80, 100, 500, 1000)
cols <- rainbow(length(levels))
contour(x, y, z, levels = levels, col = cols)
```

The following code junks are used to find extrem values

```{r}
g <- deriv(f, c("x", "y"))
g
```

Setting the gradient to $0$ and solve for $x$ and $y$ 

```{r}
crit_sol_all <- solve(g, c("x", "y"))
crit_sol_all
```

Getting rid of the variable names

```{r}
crit_sol <- crit_sol_all[1, ] %>% y_rmvars()
crit_sol
```

Converting them to an R vector

```{r}
crit <- crit_sol %>% as_r()
crit
```

To find out which type of extrem value it is, we have to check the Hessian

```{r}
H <- Hessian(f, c("x", "y"))
H
```

Inserting the roots of the gradient

```{r}
H_crit <- eval(as_r(H), list(x = crit[1], y = crit[2]))
H_crit
```

Checking for positive definiteness

```{r}
eigen(H_crit, only.values = TRUE)$values
```

Because the Hessian is positive definite, the point $[`r crit[1]`,`r crit[2]`]$ is a minimum.


# Integration Example
From Sorensen 2002 on p.17 we have

```{r}
x <- ysym("x")
f <- x * exp(-x^2/2)
f
```

```{r, results='asis'}
cat("$$", tex(f), "$$\n")
```

Now we compute the Integral

```{r}
integrate(f, x, 0, 1)
```


# Test Integration

```{r}
a <- ysym("a")
x <- ysym("x")
f <- a * x^2
f

```

Indefinite integral

```{r}
integrate(f, x)
```

The definite integral between 0 and 1

```{r}
integrate(f, x, 0, 1)
```

Try with exponential

```{r}
a <- ysym("a")
x <- ysym("x")
f <- exp(a*x)
f
```

The derivative of f

```{r}
deriv(f, x)
```

The indefinite integral

```{r}
integrate(f, x)
```

```{r}
integrate(f, x, 0, 1)
```



```{r, echo=FALSE, results='asis'}
cat(rmdhelp::get_latest_change(), "\n")
```

